---
title: "Getting Started with nitratoR"
author: "Ngoc Khanh Van Nguyen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with nitratoR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(nitratoR)
```

## Overview

The nitratoR package provides tidy monthly data from three NEON stream sites
(Arikaree, Caribou, and Lewis Run) during 12 months of 2023, including two variables about water quality, which are surface water temperature and nitrate level.
It includes:

- `nitrate_tidy`: monthly nitrate concentrations (µmol/L)

- `temp_tidy`: monthly surface-water temperature (°C)

- a helper function `monthly_median()`

- a Shiny dashboard (`launch_app()`) for interactive exploration

## Load the datasets

```{r}
data(nitrate_tidy)
data(temp_tidy)
```

## Quick function to get a single-site figure of nitrate level or water temperature

The function returns the median nitrate concentration for a given site and month.

```{r}
monthly_median(nitrate_tidy, "LEWI", "Jan")
```

## Launch the interactive app

You can visually explore both datasets interactively using the built-in Shiny dashboard:

```{r launch-app, eval=FALSE}
launch_app()
```

This app lets you choose datasets and sites, view boxplots, and explore patterns.

